---
title: "Time Series Homework Units 3 & 4"
author: "Miguel Garcia"
date: "2024-06-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1
```{r}
time_series <- c(10, 14, 20, 18, 22)

moving_average_week6 <- mean(time_series[3:5])

print(round(moving_average_week6, 2))

```

## Problem 3

```{r}
actual_value <- 50
current_forecast <- 55
alpha <- 0.2

next_forecast <- alpha * actual_value + (1 - alpha) * current_forecast

print(next_forecast)

```
## Problem 6
```{r}
library(forecast)
library(readxl)

walmart <- read_excel("C:/Users/miguel.garcia/OneDrive - Wesley International/Documents/ND/Time Series Analysis/PTSF-Datasets-20240524T164257Z-001/PTSF-Datasets/WalMartStock.xls")

fit <- Arima(walmart$Close, order=c(1,0,0))
summary(fit)

```
``` {r}
fit2 <- Arima(diff(walmart$Close, 1), order=c(1,0,0))
summary(fit2)
```
```{r}
Acf(walmart$Close, lag.max = 12, main = "")

Acf(diff(walmart$Close), lag.max = 12, main = "")

```
# Conclusion
Based on the AR(1) model results and the ACF interpretation:

The AR(1) coefficient (ar1) for the differenced series (fit2) is close to zero and not significantly different from zero. This suggests that the differenced series (increments) is consistent with white noise, characteristic of a random walk process.

Therefore, at the significance level of 0.01, the analysis indicates that the Wal-Mart closing price series is a random walk.

The Close price series is a random walk

This conclusion is based on the fact that the AR(1) model for the differenced series shows an AR(1) coefficient close to zero and not significantly different from zero, along with the expected characteristics in ACF plots for a random walk process.

## Problem 8
```{r}
library(forecast)

sales.df <- read_excel("C:/Users/miguel.garcia/OneDrive - Wesley International/Documents/ND/Time Series Analysis/PTSF-Datasets-20240524T164257Z-001/PTSF-Datasets/SouvenirSales.xls")
sales.ts <- ts(sales.df$Sales, start = c(1995, 1), end = c(2001, 12), freq = 12)

train.lm <- tslm(log(sales.ts) ~ trend + season)

forecasted_values <- forecast(train.lm, h = 1)

forecasted_value_jan_2001 <- exp(forecasted_values$mean[length(forecasted_values$mean)])

forecasted_value_jan_2001





```
## Problem 9
```{r}
library(forecast)

train.arima <- auto.arima(train.ts)

forecasted_values <- forecast(train.arima, h = 1)

forecasted_value_jan_2001 <- forecasted_values$mean[length(forecasted_values$mean)]
forecasted_value_jan_2001

```
## Problem 10
```{r}
library(forecast)

train.arima.log <- auto.arima(log(train.ts), lambda = 0)

forecasted_values <- forecast(train.arima.log, h = 1)

forecasted_value_jan_2001 <- exp(forecasted_values$mean[length(forecasted_values$mean)])

forecasted_value_jan_2001

```
## Problem 11
```{r}

actual_values_train_lm <- window(sales.ts, start = c(2000, 1), end = c(2000, 12))


forecasted_values_lm <- forecast(train.lm.B, h = 12)
forecasted_values_train_lm <- exp(forecasted_values_lm$mean)


forecasted_values_train_lm <- head(forecasted_values_train_lm, length(actual_values_train_lm))

mape_train_lm <- mean(abs((forecasted_values_train_lm - actual_values_train_lm) / actual_values_train_lm)) * 100
mape_train_lm
```

# I could not figure this error

```{r}


forecasted_values_train_arima <- forecast(train.arima, h = 12)$mean

actual_values_train_arima <- window(sales.ts, start = c(2000, 1), end = c(2000, 12))

forecasted_values_train_arima <- head(forecasted_values_train_arima, length(actual_values_train_arima))

mape_train_arima <- mean(abs((forecasted_values_train_arima - actual_values_train_arima) / actual_values_train_arima)) * 100
mape_train_arima

```
# I could not figure this error


```{r}


forecasted_values_train_arima_log <- forecast(train.arima.log, h = 12)$mean

actual_values_train_arima_log <- window(sales.ts, start = c(2000, 1), end = c(2000, 12))

forecasted_values_train_arima_log <- exp(forecasted_values_train_arima_log)
forecasted_values_train_arima_log <- head(forecasted_values_train_arima_log, length(actual_values_train_arima_log))

mape_train_arima_log <- mean(abs((forecasted_values_train_arima_log - actual_values_train_arima_log) / actual_values_train_arima_log)) * 100
mape_train_arima_log
```
## Problem 12
```{r}
lm.B <- tslm(sales.ts ~ trend + season, lambda = 0)

forecast_result <- forecast(lm.B, h = 1)

forecasted_sales_feb_2002 <- forecast_result$mean["2002-02"]
forecasted_sales_feb_2002

```



## Problem 13
```{r}
library(forecast)


log_sales.ts <- log(sales.ts)

arima.log <- auto.arima(log_sales.ts, lambda = 0)

forecast_result <- forecast(arima.log, h = 1)

forecasted_sales_feb_2002 <- exp(forecast_result$mean)["2002-02"]
forecasted_sales_feb_2002


```