---
title: "Assignment 2"
subtitle: "Machine Learning"
author: "Miguel Garcia"
output:
html_document:
toc: true
toc_float: true
theme: lumen
highlight: zenburn
df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

``` {r}
library(tidyverse)
library(rpart)
library(performanceEstimation)
library(caret)

```
# Part I: Data Preprocessing

# 1. Importing and previewing the data

```{r}
loans <- read_csv("C:/Users/miguel.garcia/Downloads/lendingclub.csv")
glimpse(loans)
```

# 2. Converting features to factors

```{r}
factor_cols <- c("Grade", "EmploymentLength", "HomeOwnership", "IncomeVerified", "LoanPurpose", "Default")
loans[factor_cols] <- lapply(loans[factor_cols], as.factor)
glimpse(loans)
```

# 3. Descriptive statistics

```{r}
summary(loans)
```

# 4. Splitting dataset into training and test sets

```{r}
set.seed(1234)
RNGkind(sample.kind = "Rounding") 
train_indices <- createDataPartition(loans$Default, p = 0.75, list = FALSE)
loans_train <- loans[train_indices, ]
loans_test <- loans[-train_indices, ]
```

# 5. Balance training set using SMOTE

```{r}
set.seed(1234)
loans_train_balanced <- smote(Default ~ ., loans_train, perc.over = 1, perc.under = 2)
loans_train_balanced %>% count(Default) %>% mutate(prop = round(n / sum(n), 4)) %>% arrange(desc(n))
```

# 6. Converting double-precision features to integers

```{r}
double_cols <- c("Delinquencies", "Inquiries", "OpenAccounts", "TotalAccounts", "PublicRecords")
loans_train_balanced[double_cols] <- lapply(loans_train_balanced[double_cols], as.integer)
loans_test[double_cols] <- lapply(loans_test[double_cols], as.integer)
glimpse(loans_train_balanced)
glimpse(loans_test)
```

# Part II: Train the Model

# 1. Training decision tree model

```{r}
model <- rpart(Default ~ .,
               method = "class",
               data = loans_train_balanced,
               cp = 0.02)

```

# 2. Visualization of the decision tree model

```{r}
library(rpart.plot)
rpart.plot(model)
```

# 3. Importance

```{r}
varImp(model)
```

# Part III: Evaluate the Model

# 1. Predicted outcomes on test dataset

```{r}
predictions <- predict(model, loans_test, type = "class")
head(predictions)
```

# 2. Adding predicted outcome column to test dataset

```{r}
loans_test$Prediction <- predictions
glimpse(loans_test)

```

# 3. Calculating accuracy

```{r}
accuracy <- mean(predictions == loans_test$Default)
accuracy
```

# Part IV: Interpret the Results

# 1. Calculating profit from loans paid back in full

```{r}
full_paid_loans <- loans_test$LoanAmount[loans_test$Default == "No"]
total_profit <- sum(full_paid_loans) * 0.12
print(total_profit)

```

# 2. Calculating loss from bad loans

```{r}
bad_loans <- loans_test$LoanAmount[loans_test$Prediction == "Yes"]
total_loss <- sum(bad_loans) * 0.70
print(total_loss)
```

# 3. Calculating total servicing cost

```{r}
servicing_cost <- sum(loans_test$LoanAmount) * 0.02
print(servicing_cost)
```

# Checking if profit or loss

```{r}
profit_or_loss <- total_profit - total_loss - servicing_cost
print(profit_or_loss)
```

# 4. Opportunity cost

```{r}
subset_full_paid_loans <- full_paid_loans[loans_test$Prediction == "Yes"]

subset_full_paid_loans_no_na <- na.omit(subset_full_paid_loans)

opportunity_cost <- sum(subset_full_paid_loans_no_na) * 0.12
print(opportunity_cost)
```