---
title: "Baseball - What Does It Take To Win?"
author: "Miguel Garcia"
format: html
server: shiny
---

## Shiny Documents

This Quarto document is made interactive using Shiny. Interactive documents allow readers to modify parameters and see the results immediately. Learn more about Shiny interactive documents at <https://quarto.org/docs/interactive/shiny/>.

## Inputs and Outputs

You can embed Shiny inputs and outputs in your document. Outputs are automatically updated whenever inputs change. This demonstrates how a standard R plot can be made interactive:

```{r}
library(dplyr)
library(caret)
library(data.table)


mlb_stats <- read.csv("C:/Users/miguel.garcia/Downloads/2019.2023.mlb.team.stats.csv")

###Data Cleaning
mlb_stats <- mlb_stats[1:150, ]
mlb_stats$mlbid <- as.character(mlb_stats$mlbid)
names(mlb_stats)[names(mlb_stats) == "R"] <- "runs_scored"
names(mlb_stats)[names(mlb_stats) == "R8"] <- "runs_allowed"
names(mlb_stats)[names(mlb_stats) == "G3"] <- "games_pitched"
names(mlb_stats)[names(mlb_stats) == "HR9"] <- "home_runs_allowed"
names(mlb_stats)[names(mlb_stats) == "strikeouts"] <- "strikeouts_pitcher"


#Run Differential column
mlb_stats$run_differential <- mlb_stats$runs_scored - mlb_stats$runs_allowed

mlb_stats$Team <- factor(mlb_stats$Team)
mlb_stats$Year <- factor(mlb_stats$Year)
mlb_stats$playoffs <- factor(mlb_stats$playoffs)

```

```{r}
filtered_champs <- mlb_stats[mlb_stats$Champs == "y", ]

selected_columns <- c("Year", "Team", "W", "L", "runs_scored", "runs_allowed", "run_differential")

print(filtered_champs[selected_columns])
```
```{r}
selected_columns <- c("Year", "HR", "SB", "BB", "H", "SO")

sums_by_year <- mlb_stats %>%
  select(all_of(selected_columns)) %>%
  group_by(Year) %>%
  summarise(across(everything(), sum, na.rm = TRUE))

print(sums_by_year)

```
```{r}
# Recode 'Champs' column to numeric
mlb_stats$Champs_numeric <- ifelse(mlb_stats$Champs == "y", 1, 0)

# Fit logistic regression model
logTest <- glm(Champs_numeric ~ HR + SB + BB + strikeouts + doubles + W + triples, data = mlb_stats, family = binomial)

summary(logTest)
##mlb_stats <- subset(mlb_stats, select = -c(Champs_numeric))

```
```{r}

mlb_stats$predictedProbs <- predict(logTest, 
                                      type = "response")


ggplot(mlb_stats, aes(strikeouts, predictedProbs)) +
  geom_line(linewidth = 1.5) +
  theme_minimal()
``` 
```{r}
twoPredictors <- glm(Champs_numeric ~ strikeouts + HR, data = mlb_stats, 
              family = binomial)

summary(twoPredictors)
```
```{r}
exp(twoPredictors$coefficients)
```
```{r}
library(plotly)


twoPredictors <- glm(Champs_numeric ~ strikeouts + HR, 
                     data = mlb_stats, 
                     family = binomial)

strikeouts <- unique(mlb_stats$strikeouts)
HR <- unique(mlb_stats$HR)

grid <- expand.grid(strikeouts, 
                    HR)

newData <- setNames(data.frame(grid), c("strikeouts", "HR"))
predictedVals <- predict(twoPredictors, 
                         newdata = newData, 
                         type = "response")

zMatrix <- matrix(predictedVals, 
            nrow = length(unique(newData$strikeouts)), 
            ncol = length(unique(newData$HR)))

plot_ly() %>% 
  add_surface(x = ~HR, y = ~strikeouts, z = ~zMatrix)
```


